---
title: "Playground"
author: "Frances Hung"
date: "10/21/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
```

## Playground


```{r}
require(gtrendsR)
require(ggplot2)
require(dplyr)
require(ggmap)
```

## Making Dataframes

This gives us a master dataframe of search frequencies of “depression” over the past 12 months in the US which relate for sure to mental health. We can take different dataframes using "$": see the dataframe for details.

```{r}
trend<-gtrends("rehab",c("US"),time="today 12-m",category=437)
trend
```

For example, this gives us search frequencies by cities in CA in the U.S.

```{r}
cities_dep<-gtrends("depression",c("US-CA"),time="today 12-m",category=437)$interest_by_city
cities_dep
```

This plots cities_dep.

```{r}
ggplot(cities_dep,aes(x=reorder(location,hits),y=hits))+geom_bar(stat="identity")+theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r}

for (i in 1:length(cities_dep$location)) {
  place=geocode(cities_dep$location[i],output="latlon",source="dsk")
  cities_dep$lat[i]=as.numeric(place[1])
  cities_dep$lon[i]=as.numeric(place[2])
}
```

```{r}
center=as.numeric(geocode("United States",source="dsk"))
mappy<-get_map(c(-119.4179,36.7783),zoom=6,scale=2,maptype = "terrain",source="google")
p=ggmap(mappy,extent="device",ylab="Latitude",xlab="Longitude")
p=p+geom_point(data=cities_dep,aes(x=lat,y=lon),size=(cities_dep$hits/50)^2)
p
```

facilities data
```{r}
facilities <- read_csv("filtered_licensed-healthcare-facility-listing-june-30-2017.csv")
facilities <- filter(facilities, LICENSE_CATEGORY_DESC == "Acute Psychiatric Hospital"|LICENSE_CATEGORY_DESC == "Psychiatric Health Facility" | LICENSE_CATEGORY_DESC =="Psychology Clinic")
View(facilities)
write.csv(facilities, file="facilities.csv")
```

suicide rate data
```{r}
suicide_rate <- read_csv("leading-causes-of-death-by-zip-code-1999-2015.csv")
colnames(suicide_rate) <- c("year", "zipcode", "causes_of_death", "count") 
suicide_rate <- filter(suicide_rate, year == "2015" & causes_of_death == "SUI")
View(suicide_rate)
write.csv(suicide_rate, file = "suicide_rate.csv")
```


